---
title: "Settlement Persistence Project, SBOM Script #7:"
subtitle: "Modeling Spatial Interactions"
author: "Rudolf Cesaretti"
date: "Last run on `r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    number_sections: true
bibliography: References.bib
csl: apa.csl
link-citations: yes
---


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

```{r, setup, include=FALSE,echo=FALSE, message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=75),tidy=TRUE)
#
rm(list = ls())
```

I do four things in this R markdown document: 

  3. Calculate settlement hierarchy metrics for AggSites at two spatial scales:
      + The global level (SBOM-region-wide) 
      + The local level (at a set radius surrounding each site)
  4. Reorganize the data and export for Script #7
  
  
# Setup 

All of the data and scripts are downloadable from the [new ASU SettlementPersist2022 github repository](https://https://github.com/rcesaret/ASUSettlementPersist2022), which can be downloaded locally as a .zip folder or cloned to your own account.

Either way, once you have done so, you will need to modify the working directory (setwd("C:/...)") path and "dir" variables in the code chunk below to match the repository location on your computer.

```{r, label='Set Local Directory Location', message=FALSE,warning=FALSE}

wd <- list()

#SET YOUR LOCAL DIRECTORY LOCATION HERE:
wd$dir <- "C:/Users/rcesaret/Dropbox (ASU)/ASUSettlementPersist2022/"
#wd$dir <- "C:/Users/TJ McMote/Dropbox (ASU)/ASUSettlementPersist2022"

wd$analysis <- paste0(wd$dir,"analysis/")
wd$data_r <- paste0(wd$dir,"data-raw/")
wd$data_p <- paste0(wd$dir,"data-processed/")
wd$data_f <- paste0(wd$dir,"data-final-outputs/")
wd$figs <- paste0(wd$dir,"figures/")
wd$funcs <- paste0(wd$dir,"functions/")

```


## Load R Packages and Custom Functions

```{r, label='Load Libraries', message=FALSE,warning=FALSE}
# Package names
packages <- c("rgdal", "rgeos", "sp", "sf", "GISTools", "raster", "Matrix", "gdistance", "lwgeom", "tidyverse", "tidyr", "classInt", "pacman", "RColorBrewer", "cowplot", "stars", "ggnewscale")#, "data.table", "zoo", "mgcv","igraph", "ggrepel","ggridges", "movecost",  "datplot", "scales",

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# load packages
invisible(lapply(packages, library, character.only = TRUE))

rm(packages,installed_packages)

#Read in custom R functions located in the wd$funcs directory folder
FUNCS <- list("splitByAttributes.R", "RS_Acoef.R", "SettHierClass_ggplot.R")
invisible(lapply(FUNCS, function(x) source(paste0(wd$funcs,x))))
rm(FUNCS)

```



## Import Data

Data we are importing:

  1. the AggSite polygon data
  2. A simple polygon calculated in QGIS that specifies a hard outter border for the catchment areas of sites (constructed for sensitivity to survey borders and sites not included in the SBOM sample)
  3. Cost-distance matrices from script #3
  4. Least cost path rasters from script #3
  5. A raster hillshade basemap for the SBOM which includes the lakes

```{r, label='Import Data', message=FALSE,warning=FALSE}

#Agg Site and catchment polygon data
All_AggPoly <- readOGR(paste0(wd$data_p,"SBOM_AggSitePoly6.gpkg"))
All_CatchPoly <- readOGR(paste0(wd$data_p,"SBOM_CatchPoly6.gpkg"))
#reorder spatial points dataframe by period
All_AggPoly <- All_AggPoly[order(All_AggPoly$PeriodNum),]
All_CatchPoly <- All_CatchPoly[order(All_CatchPoly$PeriodNum),]

# Split polygons by Phase, saved as list of SPDF
Poly_List <- splitByAttributes(spdata = All_AggPoly, attr = "Period", suffix="_SitePoly") 
Catch_List <- splitByAttributes(spdata = All_CatchPoly, attr = "Period", suffix="_CatchPoly") 

# convert spatial polygons dataframe to spatial points dataframe
coor = All_AggPoly@data[,c("East","North")] #create separate dataframe of coordinates
rownames(coor) <- as.numeric(rownames(coor)) #make sure rownames match
All_AggPts <- SpatialPointsDataFrame(coor, All_AggPoly@data, match.ID = TRUE) #convert to points
proj4string(All_AggPts) <- CRS("+proj=utm +zone=14 +datum=NAD83 +units=m +no_defs") #set CRS
All_AggPts = as(st_make_valid(st_as_sf(All_AggPts)), "Spatial") #make sure geometry is valid
All_AggPts <- All_AggPts[order(All_AggPts$PeriodNum),]#reorder by period

# Split points by Phase, saved as list of spatial points dataframes
Pts_List <- splitByAttributes(spdata = All_AggPts, attr = "Period", suffix="_Pts")

#Catchment boundary limit polygon
CatchLims <- readOGR(paste0(wd$data_r,"CatchLims.gpkg"))
CatchLims.sf = st_as_sf(CatchLims) #for ggplot 

## Hillshade Basemap Raster with lake
HillshadeLake <- raster(paste0(wd$data_r, "HillshadeLake.tif"))
HillshadeLake <- rast(HillshadeLake, crs = 26914)
Hillshade.s <- st_as_stars(HillshadeLake) #for ggplot basemap

#Cost-distance matrices
CD.mats <- readRDS(file=paste0(wd$data_p,"CDmats_list.RData"))
TrsprtNet_CDmatList <- readRDS(file=paste0(wd$data_p,"TrsprtNet_CDmatList.RData"))
```


# Modelling Interactions

Model different theoretical dimensions of interaction
--using well-established covariates and proxies
--using high-performing models 



Radiation Model 
--to model the distance decay function for different spatial interactions

Spatial Interaction Models
--Retail Model
--Doubly Constrained



LQs or similar of

--Primary Sector Trade
----AG staple
----AG type trade
----Naural resources trade
--Urban productivity specialization
----Popdens, Agglom Index, Scaling Residuals
----Transp Network centrality
----Pop Rural/Urbanization vars
----




## Radiation Model

In 2012 Filippo Simini and colleagues ([Simini et al. 2012](https://dspace.mit.edu/handle/1721.1/77896)) presented a new model, designed specifically to model human geographic mobility called the radiation model. This model was created explicitly as an alternative to various gravity models and, in certain cases, was demonstrated to generate improved empirical predictions of human population movement between origins and destinations. This model shares some basic features with gravity models but importantly, the approach includes no parameters at all. Instead, this model uses simply measures of population at a set of sites and the distances between them. That is all that is required so this model is relatively simple and could likely be applied in many archaeological cases. Let's take a look at how it is formally defined: 

$$T_{ij} = T_i \frac{m_in_j}{(m_i + s_{ij})(m_i + n_j + s_{ij})}$$

where

* $T_i$ is the total number of "commuters" or migrating individuals from node $i$.
* $m_i$ and $n_j$ are the population estimates of nodes $i$ and $j$ respectively.
* $s_{ij}$ is the total population in a circle centered at node $i$ and touching node $j$ excluding the populations of both $i$ and $j$.

We have defined a function for calculating radiation among a set of sites using just two inputs:

* **`pop`** is a vector of population values.
* **`d_mat`** is a distance matrix among all nodes.


```{r}

```


## Rihll-Wilson Retail Model

```{r}

```




# Recombining and Reorganizing the Data

```{r, 'Recombining the Data Part 2', message=FALSE, warning=FALSE}
# Convert lists of period-wise sites/catchments to single SPDF objects
All_Agg_SitePoly <-  do.call(rbind, Poly_List)
All_Agg_CatchPoly <- do.call(rbind, Catch_List)

# variables from site data that needs transfer over to catchment areas
colz1 = setdiff(colnames(All_Agg_SitePoly@data),colnames(All_Agg_CatchPoly@data))
# variables from catchment areas that needs transfer over to site data
colz2 = setdiff(colnames(All_Agg_CatchPoly@data),colnames(All_Agg_SitePoly@data))

#reorder the data to match
All_Agg_SitePoly <- All_Agg_SitePoly[order(All_Agg_SitePoly$AggSite),]
All_Agg_CatchPoly <- All_Agg_CatchPoly[order(All_Agg_CatchPoly$AggSite),]

#check to see that the two datasets are in the right order
#identical(All_Agg_SitePoly@data$AggSite, All_Agg_CatchPoly@data$AggSite)

# Site data to catchment areas
Site_to_Catch <- All_Agg_SitePoly@data %>% select(!!!syms(colz1))
All_Agg_CatchPoly@data <- cbind(All_Agg_CatchPoly@data,Site_to_Catch)

# catchment area data to sites
Catch_to_Site <- All_Agg_CatchPoly@data %>% select(!!!syms(colz2))
All_Agg_SitePoly@data <- cbind(All_Agg_SitePoly@data,Catch_to_Site)

#Reorganize the data

ordering <- c(
  #ID VARIABLES
      "AggSite","AggID","Site","East","North","SurvReg","Number","CerPhase","Period", 
      "PeriodType","PeriodLength","PeriodNum","PeriodBegin","PeriodEnd", 
      "OccSeqLoc","OccSeqLoc.Sites","SubOccSeqLoc","SubOccSeqLoc.Sites",
      "ComponentNum", "ComponentSites",
  #CHRONOLOGICAL VARIABLES
      "PeriodInterval", "PeriodBegin", "PeriodBegin.era", "PeriodMidpoint", 
      "PeriodMidpoint.era", "PeriodEnd", "PeriodEnd.era", "PeriodLength",
  #OCCUPATION VARIABLES (COUNTS)
      "Occ.EF","Occ.EF_MF","Occ.MF","Occ.MF_LF","Occ.LF","Occ.LF_TF",
      "Occ.TF","Occ.TF_CL","Occ.CL","Occ.CL_ET","Occ.ET","Occ.ET_LTAzI", 
      "Occ.LTAzI","Occ.LTAzI_EA","Occ.EA","Occ.EA_LA","Occ.LA","Occ.TOT",
  #SUBOCCUPATION VARIABLES (COUNTS)
      "SubOcc.EF","SubOcc.EF_MF","SubOcc.MF","SubOcc.MF_LF","SubOcc.LF",
      "SubOcc.LF_TF","SubOcc.TF","SubOcc.TF_CL","SubOcc.CL","SubOcc.CL_ET",
      "SubOcc.ET","SubOcc.ET_LTAzI","SubOcc.LTAzI","SubOcc.LTAzI_EA",
      "SubOcc.EA","SubOcc.EA_LA","SubOcc.LA","SubOcc.TOT",
  #SITE AREA AND OCCUPATIONAL DENSITY VARS
      "Area_ha","Perim_m2","SherdDens","Tot.Assemb","FwOvlp.Assemb", 
      "BwOvlp.Assemb", "Net.Assemb",
  #STEP #2 DEMOGRAPHIC VARIABLES
      "Population.s2","Log_Population.s2", "ApportAssemb", "PopDens.s2",
      "UrbanScale.s2", "UrbanPop.s2","RuralPop.s2", "PctUrban.s2","PctRural.s2",
      "Prior", "Observed", "MeanOccuProb", 
  #STEP #2 DEMOGRAPHIC RATES
      "Pct_deltaPop12", "Pct_deltaPop01", "r12_Pert", "r01_Pert",
      "r23_Pert", "rPert", "rPert0", "rPert2", 
      "Pct_UrbdeltaPop12", "Pct_UrbdeltaPop01", "Urb_r12_Pert", "Urb_r01_Pert",
      "Urb_r23_Pert", "Urb_rPert", "Urb_rPert0", "Urb_rPert2",
      "Abs_rPert","RS_rPert","LogRS_rPert","LogAbs_rPert","Q.rPert","Q.Abs_rPert",
      "Q.RS_rPert","Q.LogRS_rPert","Q.LogAbs_rPert","Urb_Abs_rPert",
      "Urb_RS_rPert","Urb_LogRS_rPert","Urb_LogAbs_rPert","Q.Urb_rPert","Q.Urb_Abs_rPert",
      "Q.Urb_RS_rPert","Q.Urb_LogRS_rPert","Q.Urb_LogAbs_rPert",
      "Qp.rPert","Qp.Abs_rPert","Qp.RS_rPert","Qp.LogRS_rPert","Qp.LogAbs_rPert",
  #STEP #1 DEMOGRAPHIC VARIABLES
      "Population.s1","PopDens.s1","UrbanScale.s1","UrbanPop.s1","RuralPop.s1", 
      "PctUrban.s1","PctRural.s1",
  #CONTINUITY VARIABLES
      "AreaBwCont","AreaFwCont","PopBwCont","PopFwCont","FwOvlp.Sites",
      "FwOvlp.Area","FwOvlp.Pop","BwOvlp.Sites","BwOvlp.Area","BwOvlp.Pop",
  #PERSISTENCE VARIABLES
      "Found","FoundInit","Abandon","Persist","DewarType",
      "OccuTime","OccuInertia","UrbOccuTime","UrbOccuInertia",
  #STEP #4 TRANSPORT NETWORK VARIABLES
      "TranspDens.pct", "TranspDens.rank","centr_deg","centr_btw","centr_eig",
      "centr_clos","centr_hrmo","centr_hub","centr_auth","centr_pgrk",
      "centr_deg_n","centr_btw_n","centr_eig_n","centr_clos_n","centr_hrmo_n",
      "centr_hub_n","centr_auth_n","centr_avg","trans_loc","trans_locavg",
      "density","connectiv","trans_glob","cntrlz_deg","cntrlz_btw","cntrlz_eig",
      "cntrlz_clo","cntrlz_hub","cntrlz_auth","cntrlz_deg_n","cntrlz_btw_n",
      "cntrlz_eig_n","cntrlz_clo_n","cntrlz_hub_n","cntrlz_auth_n","cntrlz_avg",
  #STEP #4 CATCHMENT AREA AND POP DENSITY VARIABLES
      "Catchment_ha", "CatchmentBeyond_ha", "Catch_Popdens", 
      "Catch_Popdens_PropMax","Catch_Popdens_Rank","CatchB_Popdens",
      "CatchB_Popdens_PropMax", "CatchB_Popdens_Rank", 
  #STEP #5 CATCHMENT ENVIRONMENT/TOPOGRAPHY VARIABLES
      "NPP.tot","NPP.avg","EZ.avg","EZ.sd","TRI.tot","TRI.avg","IrrigPot.tot",
      "IrrigPot.avg","WetAgPot.tot","WetAgPot.avg","IntnsCost.tot",
      "IntnsCost.avg","IntnsCostNW.tot","IntnsCostNW.avg","AGPot.tot",
      "AGPot.avg","AGPotNW.tot","AGPotNW.avg","PopPressure","PopPressureNW",
      "ErosionPot.tot","ErosionPot.avg", 
  #STEP #6 SETTLEMENT HIERARCHY DEMOG VARS
      "SetHierLevel","ScalResid","ScalSlope","ScalInt","Pop_PropMax","Pop_Rank",
      "UrbanPop_PropMax","UrbanPop_Rank","PopDens_PropMax","PopDens_Rank",
      "UrbanScale_PropMax","UrbanScale_Rank",
  #STEP #6 SETTLEMENT HIERARCHY/DEMOG LOCATION QUOTIENTS
      "LQ_UrbPop","FLQ_UrbPop","LQ_UrbOccuTime","FLQ_UrbOccuTime","LQ_UrbOccuInertia",
      "FLQ_UrbOccuInertia","LQ_PopFwCont","LQ_PopBwCont","LQ_AreaFwCont","LQ_AreaBwCont",
      #"LQ_Pct_UrbdeltaPop12","FLQ_Pct_UrbdeltaPop12","LQ_PopDens",
      #"FLQ_PopDens","LQ_Catch_Popdens","FLQ_Catch_Popdens","LQ_Urb_Abs_rPert",
      #"FLQ_Urb_Abs_rPert","LQ_Q.Urb_rPert","FLQ_Q.Urb_rPert",
  #STEP #6 SETTLEMENT HIERARCHY/DEMOG SPATIAL GINIS
      "spGini_Pop","spGini_UrbPop","spGini_OccuTime","spGini_UrbOccuTime",
      "spGini_OccuInertia","spGini_UrbOccuInertia","spGini_Pct_deltaPop12",
      'spGini_Pct_UrbdeltaPop12',"spGini_Area_ha","spGini_Catchment_ha",
      "spGini_Abs_rPert","spGini_Urb_Abs_rPert","spGini_Q.rPert","spGini_Q.Urb_rPert",
      "spGini_FwOvlp.Pop","spGini_BwOvlp.Pop","spGini_FwOvlp.Area","spGini_BwOvlp.Area",
      "spGini_PopPressure",
  #STEP #6 TRANSPORT NETWORK SPATIAL GINIS
      "spGini_centr_avg","spGini_centr_deg","spGini_centr_btw",
      "spGini_centr_eig","spGini_centr_clos","spGini_centr_hrmo","spGini_centr_hub",
      "spGini_centr_auth","spGini_centr_pgrk","spGini_trans_loc","spGini_SetHierLevel",
  #SURVEY METADATA
      "M_Sites","M_SiteCode","M_SiteName","M_FieldSite.Region",
      "M_FieldSite.Period","M_SurveyYearNumber","M_Supervisor","M_Map",
  #OLD tDAR BOM SURVEY VARIABLES
      "O_Elev","O_ElevMed","O_ElevMin","O_ElevMax","O_EZcode",
      "O_EnvironmentalZone","O_Soil","O_SoilMed","O_SoilMin","O_SoilMax",
      "O_Erosion","O_ErosionMed","O_ErosionMin","O_ErosionMax","O_ModernUse",
      "O_ModernSettlement","O_Rainfall","O_Area","O_MoundDomestic",
      "O_MoundCeremonial","O_MoundQuestionable","O_MoundTotal",
      "O_MoundRecorded","O_DMoundArea","O_Architecture","O_TerraceConfidence",
      "O_TerraceExtent","O_Sherd","O_SherdMed","O_SherdMin","O_SherdMax",
      "O_Rubble","O_RubbleMed","O_RubbleMin","O_RubbleMax","O_Population",
      "O_PopMin","O_PopMax","O_PopMethod","O_stcode","O_SiteType",
      "O_SubPeriod1","O_SubPeriod2","O_OccEF","O_OccMF","O_OccLF","O_OccTF",
      "O_OccCL","O_OccEC","O_OccMC","O_OccLC","O_OccET","O_OccLT","O_OccAZ",
      "O_OccEA","O_OccLA","O_OccTot","O_OccSeqLoc","O_SubOc1","O_SubOc2",
      "O_PdDupSite","O_Group","O_Comments") 

#Make sure everything is kosher
#setdiff(colnames(All_Agg_CatchPoly@data),ordering)
#setdiff(colnames(All_Agg_SitePoly@data),ordering)
#setdiff(ordering,colnames(All_Agg_CatchPoly@data))
#setdiff(ordering,colnames(All_Agg_SitePoly@data))

# Reorder the data for both sites and catchment areas
All_Agg_SitePoly@data <- All_Agg_SitePoly@data %>% select(!!!syms(ordering))
All_Agg_CatchPoly@data <- All_Agg_CatchPoly@data %>% select(!!!syms(ordering))
```


# Export Data for Script #8

```{r, 'Export Data for Step #8', message=FALSE,warning=FALSE}

#AggSite polygons
writeOGR(All_Agg_SitePoly, paste0(dir2,"SBOM_AggSitePoly7.gpkg"), "SBOM_AggSitePoly7", driver = "GPKG", overwrite_layer=TRUE)

#Catchment areas
writeOGR(All_Agg_CatchPoly, paste0(dir2,"SBOM_CatchPoly7.gpkg"), "SBOM_CatchPoly7", driver = "GPKG", overwrite_layer=TRUE)

```


# References
